\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Wed. Sept 20}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2017} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 3}}
\date{Start this assignment before Sunday night!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 1-6, 1-7, Chapter 1}
\item{Handout 3}
\item{Lecture Notes}
\end{itemize}

\section*{ Do: }
\paragraph{A: Problems on Reviewing of Orthogonal transformations, 
Rotations, Reflections and Rigid Motions
  in $R^n$.}
\begin{itemize}

{\item a) Let $\rho$ and $\tau$ be two orthogonal transformations on
  an Euclidean space $(V^n, <,>)$.  Prove that the composition of
  $\rho$ and $\tau$ is again an orthogonal transformation of $(V^n,
  <,>)$.  So is the inverse of $\rho$.}
 
{\item b) Problem 6 on page 23, Section 1-5, Baby Do Carmo.}
\end{itemize}

\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
{\item a) Show SO(n) is a group with respect to the usual matrix 
multiplication. (Later, we will see that SO(n) is in fact a Lie group.)}
{\item b) Show that the mirror reflection $\tau$ (as defined in the 
lecture) is an orthogonal transformation and $\tau^2 = id$, where 
$id$ is the identity transformation. }
\end{itemize}


\paragraph{C: Other Problems}

\

\noindent {\bf Choose 2 problems out of following problems:}
\begin{itemize}
{\item a) Problem 3 on page 7, Section 1-3, Baby Do Carmo.}
{\item b) Problem 5 on page 8, Section 1-3, Baby Do Carmo.}
{\item c) Problem 6 on page 8, Section 1-3, Baby Do Carmo.}
\end{itemize}


\noindent {\bf Choose 3 problems out of following problems:}
\begin{itemize}
{\item a) Problem 1 on page 22, Section 1-5, Baby Do Carmo.}
{\item b) Problem 2 on page 22, Section 1-5, Baby Do Carmo.}
{\item c) Problem 5 on page 23, Section 1-5, Baby Do Carmo.}
{\item d) Problem 12 on page 25, Section 1-5, Baby Do Carmo.}
\end{itemize}

\paragraph{D: Extra Credit Problems}
\begin{itemize}
{\item Problems 7, 8  on page 22-23, Section 1-5, Baby Do Carmo.}
\end{itemize}


\end{document}







