\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Monday November 29, 1:15 PM}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2004} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 10}}
\date{Start this assignment before Sunday night!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 4-3, 4-4 of Chapter 4, Appendix of Chapter 5 starting on page 456,
and Section 5-2 of Chapter 5.}
\item{ Read and fill out the blanks for the most recent handouts. }
\item{Lecture Notes.}
\end{itemize}

\section*{ Do: }
\paragraph{A: Problems on Reviewing of Point-Set Topology of 
Euclidean Spaces.}

\begin{itemize}
{\item a) What do we mean by a set is open, closed, bounded,  
compact, or connected? Give definitions.}

{\item b) Find all possible ``if and only if'' conditions to 
characterize a set being compact.}

\end{itemize}

\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
{\item a) Carry out all the details in finding the Christoffel symbols
  of a surface of revolution.} 
%{\item b) We learned in the class that if a regular surface $S$ is 
%compact, connected, and with constant Gaussian curvature. Then $S$ 
%must be a sphere.  Give examples to show that we can not drop the condition 
%that $S$ being compact or $S$ being connected. (If you use the
%  example on the handout, please carry out the all details.)  Note:
%  The example there looks like a piece a torus, in fact, it is NOT.}   
\end{itemize}

\paragraph{C: Other Problems}
\begin{itemize}

{\item a) Problem 1 on page 237, Section 4-3, Baby Do Carmo.}
{\item b) Problem 3 on page 237, Section 4-3, Baby Do Carmo.}
{\item c) Problem 4 on page 237,  Section 4-3, Baby Do Carmo.}
{\item d) Problem 6 on page 237,  Section 4-3 Baby Do Carmo.}
{\item e) Problem 8 on page 237,  Section 4-3, Baby Do Carmo.}
{\item f) Problem 9 on page 237,  Section 4-3, Baby Do Carmo.}
\end{itemize}


\paragraph{E: Reminder}
\begin{itemize}
\item{\bf Problem Sessions: Sunday 8:30--9:30 PM, BK 134.}
\end{itemize}


\end{document}









