\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Wed. October 4}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2017} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 5}}
\date{Start this assignment before Sunday night!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 2-2, 2-3, 2-4 and  Appendix (starting on page 118) on A Brief
Review of Continuity and Differentiability}
\item{Handouts 6 and 7}
\item{Lecture Notes}
\end{itemize}

\section*{ Do: }
\paragraph{A: Problems on Reviewing of Continuity and Differentiability}
\begin{itemize}
{\item a) Prove the proposition 7 on page 127, Baby Do Carmo.} 
{\item b) Prove the proposition 8 on page 129, Baby Do Carmo.}
{\item c) Rewrite Example 11 on page 132 of Baby Do Carmo and explain 
clearly why the Inverse Function Theorem (page 131) is true only in a
neighborhood of a point $p$. } 
{\item d) Show that an infinite cylinder after deleting a vertical line is 
diffeomorphic to a plane.}
\end{itemize}

\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
{\item a) Use Inverse Function Theorem to give a proof 
of proposition 2, page 59, Baby Co Carmo.}

{\item b) Use Inverse Function Theorem to give a proof 
of proposition 4, page 64, Baby Co Carmo.}
\end{itemize}

\paragraph{C: Other Problems}
\begin{itemize}
{\item a) Problem 7 on page 66, Section 2-2, Baby Do Carmo.}
{\item b) Problem 11 on page 66, Section 2-2, Baby Do Carmo.}
{\item c) Problem 1 on page 80, Section 2-3, Baby Do Carmo.}
{\item d) Problem 8 on page 80, Section 2-3, Baby Do Carmo.}
{\item e) Problem 10 on page 81, Section 2-3, Baby Do Carmo.}
{\item f) Problem 12 on page 81, Section 2-3, Baby Do Carmo.}
{\item g) Problem 15 on page 82, Section 2-3, Baby Do Carmo.}
\end{itemize}

\paragraph{D: Extra Credit Problems}
\begin{itemize}
{\item Problem 13 on page 82, Section 2-3, Baby Do Carmo.}
\end{itemize}

\end{document}








