\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Wednesday, December 6,  1:15PM}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2004} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 11}}
\date{Start this assignment early!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 4-4, 4-5 of Chapter 4.}
\item{ Read and fill out the blanks for the most recent handouts.}
\item{Lecture Notes.}
\end{itemize}

\section*{ Do: }
\paragraph{A: Problems on Reviewing Directional and Covariant
Derivatives on $E^3$.}

\begin{itemize}
{\item a) Prove (ii) and (iii) of the theorem on page 5, Handout 19.} 
\end{itemize}

\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
{\item a) Carry out all the details in study locally the geodesics 
of a surface of revolution.} 
{\item b) Show that any geodesics of a paraboloid of revolution 
$z = x^2 + y^2$ which is not a meridian intersects itself an infinite 
number of times (See Example 6, page 258).}   
\end{itemize}

\paragraph{C: Other Problems}
\begin{itemize}

{\item a) Problem 1 on page 260, Section 4-4, Baby Do Carmo.}
{\item b) Problem 3 on page 260, Section 4-4, Baby Do Carmo.}
{\item c) Problem 4 on page 260, Section 4-4, Baby Do Carmo.}
{\item d) Problem 5 on page 260,  Section 4-4 Baby Do Carmo.}
{\item e) Problem 9 on page 261,  Section 4-4, Baby Do Carmo.}
{\item f) Problem 10 on page 261,  Section 4-4, Baby Do Carmo.}
{\item g) Problem 2 on page 282,  Section 4-5, Baby Do Carmo.}
{\item h) Problem 4 on page 282,  Section 4-5, Baby Do Carmo.}
\end{itemize}


%\paragraph{E: Reminder}
%\begin{itemize}
%\item{\bf Problem Sessions: Sunday 9:00--10:00PM, BK 134 and
%  Office Hours: Mon 3:00PM--4:00PM and Tuesday 2:30PM--3:30PM or 
%by an appointment.  Please feel free 
%for help!  I am here for YOU!}
%\end{itemize}


\end{document}









